---
title: Effect of different water types of freshwater copepod survival
date: "Current as of `r Sys.Date()`"
output: 
  html_document:
          code_folding: hide
          code_download: true
          toc: true
          toc_float: true
  github_document:
          html_preview: false
          toc: true
          toc_depth: 3
---

```{r setup, include=T, message = F, warning = F, echo = F}
knitr::opts_chunk$set(
  echo = knitr::is_html_output(),
  fig.align = "center",
  fig.path = "../Figures/markdown/",
  dev = c("png", "pdf"),
  message = FALSE,
  warning = FALSE,
  collapse = T
)

theme_matt = function(base_size = 18,
                      dark_text = "grey20"){
  mid_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[2]
  light_text <-  monochromeR::generate_palette(dark_text, "go_lighter", n_colours = 5)[3]
  
  ggpubr::theme_pubr(base_family="sans") %+replace% 
    theme(
      panel.background  = element_rect(fill="transparent", colour=NA), 
      plot.background = element_rect(fill="transparent", colour=NA), 
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key = element_rect(fill="transparent", colour=NA),
      text = element_text(colour = mid_text, lineheight = 1.1),
      title = element_text(size = base_size * 1.5,
                           colour = dark_text),
      axis.text = element_text(size = base_size,
                               colour = mid_text),
      axis.title = element_text(size = base_size * 1.2),
      axis.title.y = element_text(angle = 90,
                                  vjust = 4,
                                  margin = margin(t = 0, r = 0, b = 0, l = 10)),
      axis.title.x = element_text(vjust = -1,
                                  margin = margin(t = 0, r = 0, b = 10, l = 0)),
      legend.text = element_text(size=base_size * 0.9),
      legend.title = element_text(size = base_size * 0.9, 
                                  face = "bold"),
      plot.margin = margin(0.25, 0.25, 0.25, 0.25,"cm")
    )
}
```

Copepods were collected from Lake Champlain on April 26th 2023. Water temperature was ~7˚C. Several species were abundant, but we only examined survival of adult female *Leptodiaptomus sicilis* here. We expect results will be broadly generalizable to other species

# Survival
We examined survival in three different water types: 63-um filtered lake water, bottled spring water (Crystal Geyser brand), and reverse-osmosis filtered water. Survival was tracked for `r length(unique(surv_data$ID))` individual copepods. Shown below is the change in survival over time. Lines for the different treatments are shown in different colors. Data for the two groups is lumped together.  

```{r surv-curves}
mort_data = mort_data %>%  
  mutate("status" = if_else(death_day <= max(mort_data$death_day) - 1, 1, 0),
         treatment = factor(treatment, levels = c("FLW","BSW","ROW")),
         exp_rep = factor(exp_rep, levels = c("W","F"))
  )

surv_model = survfit(Surv(death_day, status) ~ treatment, data = mort_data)

#summary(surv_model, times = c(1:14))

ggsurvplot(
  fit = surv_model,
  xlab = "Days",
  ylab = "Survival Proportion",
  title = "Kaplan-Meier Proportional Survival Estimates",
  conf.int = TRUE,
  pval = FALSE)
```

Shown below are estimates of the proportional hazard ratio (the change in risk of mortality relative to the control) for the different treatments, but also for the two experimental replicates. 
```{r hazard-ratios}

cox_model <- coxph(Surv(death_day, status) ~ treatment + exp_rep + volume, data = mort_data)

#summary(cox_model)

ggforest(cox_model, data = mort_data)

```

# Body Size 
After 20 days, surviving females were measured (prosome length; mm). Sizes ranged from `r paste(min(size_data$length), " mm to ", max(size_data$length), " mm", sep = "")`.

```{r size-hist}
ggplot(size_data, aes(x = length)) + 
  geom_histogram(binwidth = 0.025, colour = "black") + 
  labs(x = "Prosome Length (mm)") + 
  theme_matt()
```

# Fecundity
*Leptodiaptomus* retains eggs in loosely attached ovisacs, which allows us to measure fecundity. Here we refer to fecundity as the number of eggs contained in the clutch. 

The initial clutch size (reflecting fecundity in the field) varied between `r paste(min(filter(clutch_data, clutch_num == 1)$clutch_size), "and", max(filter(clutch_data, clutch_num == 1)$clutch_size), sep = " ")` eggs. 

```{r clutch-hist}
clutch_data %>% 
  filter(clutch_num == 1) %>%  
  ggplot(aes(x = clutch_size)) +
  geom_histogram(binwidth = 2, fill = "grey90", colour = "grey20") + 
  labs(x = "Fecundity (number of eggs)") + 
  theme_matt()
```

```{r clutch-treatments, include = F}
if(length(unique(clutch_data$clutch_num)) > 1){
  ggplot(clutch_data, aes(x = clutch_num, y = clutch_size, group = ID, colour = treatment)) + 
    geom_point(size = 2) + 
    geom_line(linewidth = 2) + 
    scale_x_continuous(breaks = seq(from = 1, to = max(clutch_data$clutch_num))) + 
    labs(x = "Clutch Number", 
         y = "Fecundity (number of eggs)") + 
    theme_matt()
}else{
  ggplot(clutch_data, aes(x = treatment, y = clutch_size)) + 
    geom_boxplot(outlier.colour = NA) + 
    geom_point(size = 3, alpha = 0.4, position = position_jitter(width = 0.05, height = 0)) + 
    labs(x = "Treatment", 
         y = "Fecundity (number of eggs)") + 
    theme_matt()
}

```

Eggs took between `r paste(min(filter(clutch_data, clutch_num == 1)$clutch_hold_time, na.rm = T), "and", max(filter(clutch_data, clutch_num == 1)$clutch_hold_time, na.rm = T), sep = " ")` days to hatch. 
```{r hold-times}
ggplot(clutch_data, aes(x = treatment, y = clutch_hold_time)) + 
  geom_boxplot() + 
  geom_point(size = 3, alpha = 0.4, position = position_jitter(width = 0.05, height = 0)) + 
  ylim(c(0, max(clutch_data$clutch_hold_time, na.rm = T))) + 
  labs(x = "Water Treatment", 
       y = "Clutch Hold Time (days)") + 
  theme_matt()
```

```{r size-model}
size_fec.model = lm(data = size_egg, clutch_size ~ length)

egg_slope = coefficients(size_fec.model)[2]
```

We might expect that larger females produce larger clutches of eggs. While limited by the small sample size, we see a positive trend here, where fecundity increases by about `r round(unname(egg_slope), digits = 1)` eggs per 0.1 mm in size.

```{r size-fecundity-plot}
size_egg = inner_join(clutch_data, mutate(size_data, cup = as.numeric(cup)), by = c("exp_rep", "cup", "treatment", "volume")) %>% 
  select(exp_rep, cup, treatment, clutch_size, length)

ggplot(size_egg, aes(x = length, y = clutch_size)) + 
  geom_smooth(method = "lm", size = 2, colour = "grey70") + 
  geom_point(size = 4, aes(colour = treatment)) + 
  scale_colour_manual(values = c("BSW" = "turquoise3", "FLW" = "darkolivegreen3")) + 
  labs(x = "Prosome Length (mm)", 
       y = "Clutch Size (# eggs)") + 
  theme_matt()
```

```{r size-fecundity-anova}
knitr::kable(car::Anova(size_fec.model, type = "III"))
```


